\documentclass{article}

%% Page Margins %%
\usepackage{geometry}
\geometry{
    top = 0.75in,
    bottom = 0.75in,
    right = 0.75in,
    left = 0.75in,
}

\usepackage{graphicx}
\usepackage{hyperref}

\title{CSC209 Mini-Project - 4-Band Audio Equalizer}

\author{Name 1 & Name 2}

\begin{document}
\maketitle

\section{Project Overview}\label{desc}

Hello! Our CSC209 project is a 4-band audio equalizer. This program takes in a \texttt{.wav} file and returns a new \texttt{.wav} file with modified amplitudes in different frequency ranges (bands).

\subsection{A Brief Explanation of Manipulating Audio Data from \texttt{.wav} Files}\label{wav}

One of the more time-consuming parts of this project was understanding how \texttt{.wav} audio data was stored and how we would use it to fulfil our goal with this project. Information on the topic was also surprisingly hard to come by, which is why we include this short summary of the what needed to understand the how of our program. \\

\texttt{.wav} files are separated into two main parts: the metadata (44 bytes long) which contains information about the file, and the audio data itself (the remaining bytes in the file), called \textbf{Pulse Code Modulation Data} (or PCM data for short).

PCM data is stored as a sequence of integers in binary, of which the size depends on a parameter called \textbf{bit depth}. Common values for bit-depth are 16 bits (CD quality), 24, and 32 bits (professional recordings). PCM data effectively acts as samples of audio, or snippets of the wave form of the audio. A \texttt{.wav} file stores a certain number of samples per second, usually at a frequency of $44.1$Khz (if working with audio) or $48$Khz (if working with video). This value is called the \textbf{sample rate}.\\

Wave forms are great, but not quite what we want for this project; we are looking for \textit{frequencies}.

To get frequencies from wave forms, we must convert the time domain into the frequency domain. There are three steps to this process. First, we must normalize the wave forms so that their values are all between $0$ and $1$. Then, we must use a \textbf{windowing function} which will simplify the wave forms so that their ends are flat. Finally, the conversion itself is done using \textbf{Fast Fourier Transforms (FFT)}. These transforms take a certain number of wave forms (usually a power of $2$ for algorithm efficiency) and converts them into frequency bins ranging from $0$Hz to the \textbf{Nyquist frequency}, which is equal to the sample rate divided by two. You can imagine the frequency bins as a "snapshot" of how loud the audio is at different pitches.\\

These snapshots are exactly what we are looking for! But once we are done with amplifying them however we want, how do we convert them back to wave forms? The answer to this question is to go through the same steps in reverse order!

We will first want to do an \textbf{Inverse Fast Fourier Transform (IFFT)} on each snapshot, and then do an \textbf{inverse windowing function} to convert all our snapshots back into wave forms. It is then just a matter of putting them back together in a file with metadata, and our audio has been modified!

\subsection{Application Processing Steps}\label{steps}

The amplification process is done through multiple steps, listed below. In our case, the window function we are using is the Hann Window Function.

\begin{enumerate}
    \item Fetch metadata and PCM data from the \texttt{.wav} file input.
    \item Normalize PCM data and apply Hann Window Function on slices of $2048$ PCM data integers.
    \item Perform FFT on each slice of $2048$ integers.
    \item Amplify specific frequency ranges by a given amount.
    \item Perform IFFT on each slice.
    \item Apply Inverse Hann Window Function on each slice.
    \item Create new \texttt{.wav} file with the same metadata as the original file.
    \item Append modified data to it and return new file.
\end{enumerate}

Our project belongs to Category 1: Multi-Process Application Using Pipes, as the structure of audio and the goal of our program fits this wonderfully. Indeed, most audio is in stereo format, meaning that the left and right side (channels) each have their own data, so this already calls for multi-process computations, one for each side. Then, we also decided to perform steps 3-5 in separate children processes to the processes for each audio channel. We are doing this by having a worker pool of 20 workers per channel, for a total of 40 workers. The details regarding relations between these processes is detailed in \hyperref[arch]{Section 3}. Their communications are explained in \hyperref[comm]{Section 4}.

\section{Build and Run Instructions}\label{build}
Our application comes bundled with a \texttt{Makefile}. Running \texttt{make} creates the executable under the name \texttt{eq}. This process also creates \texttt{.o} files, which can be removed using \texttt{make clean}.\\

\texttt{eq} takes in the file name of the \texttt{.wav} file to modify, and will return a new file named \texttt{modified.wav} once it is done and does not encounter any errors. If \texttt{modified.wav} already exists before running the program, it will be overwritten. The file must be a \texttt{.wav} file, and must have a 16-bit PCM format. These will be checked in the program when the file is read before performing any data manipulation.

\newpage
\section{Architecture Diagram}\label{arch}

\newpage
\section{Communication Protocol}\label{comm}

\section{Concurrency Model}\label{conc}

\section{Error Handling and Robustness}\label{err}

\section{Team Contributions}\label{team}

\end{document}